---
title: "fitness tracker analysis"
author: "Adithya Ajay"
date: "20/10/2020"
output:
    html_document: default
---
With the help of fitness tracker devices it has become easy and inexpensively to retirve data regarding users workouts and their quality.
In this project we will analyse the data retreived from these devices in order to improve the workout quality of users


```{r, message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(knitr)
```
OBtianing data from files saved in the system
```{r echo=TRUE}
train1 <- read.csv("pml-training.csv")
test1 <- read.csv("pml-testing.csv")
divtrain <- createDataPartition(train1$classe, p=0.6, list=FALSE)
divtrain1 <- train1[divtrain, ]
divtest1 <- train1[-divtrain, ]
```

organising and cleaning data by removing data with variance close to 0
```{r echo=TRUE}

closeto0 <- nearZeroVar(divtrain1)
divtrain1 <- divtrain1[, -closeto0]
divtest1 <- divtest1[, -closeto0]

highNA <- sapply(divtrain1, function(x) mean(is.na(x))) > 0.95
divtrain1 <- divtrain1[, highNA==F]
divtest1 <- divtest1[, highNA==F]

divtrain1 <- divtrain1, -(1:5)]
divtest1 <- divtest1[, -(1:5)]
```

After cleaning we will use machine learing models for predicting purposes:

```{r echo=TRUE}
modFit <- randomForest(classe ~ ., data=myTraining)
modFit

predict <- predict(modFit, myTesting, type="class")
confusionMatrix(myTesting$classe, predict)
```

```{r echo=TRUE}
modFit_T <- rpart(classe~., myTraining)

predict_T <- predict(modFit_T, myTesting, type="class")
confusionMatrix(myTesting$classe, predict_T)
```

```{r, message=FALSE, warning=FALSE}
control_GBM <- trainControl(method = "repeatedcv", number=5, repeats=1)
modFit_GBM <- train(classe~., myTraining, method="gbm", trControl=control_GBM, verbose=FALSE)
```
```{r echo=TRUE}

predict_GBM <- predict(modFit_GBM, myTesting)
confusionMatrix(predict_GBM, myTesting$classe)
```

From the model we find that for given set accuracy of Random forest is 99.6%, decision tree=75.4%, GBM model=98.8%
and expected sample error are as follows:
RF=0.4%
DT=24.6%
GBM=1.2%


By using the most accurate prediction model found i.e., random forest we run the algorithm for 20 test cases of testing data set

```{r echo=TRUE}
predict_test <- predict(modFit, testing_data, type = "class")
predict_test
```
